---
title: "DSCI401_Exam1"
author: "Bernard Ofori Boateng"
date: "2023-10-18"
output: html_document
---

```{rsetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# loading the package and tables
library(tidyverse)
```


# Question 1
```{r}
#Loading the dataset
employee <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI401/main/data/Employee.csv")
```


```{r}
#Reading the first few rows of the dataset
head(employee)
```

```{r}
# getting the shape of the dataframe
dim(employee)
```
```{r}
#Getting the columns
names(employee)
```
#(a)

### For each education level, what percentage of the employees are female
```{r}
#Grouping the dataframe by education and gender
Grouped <- employee %>%
  group_by(Education, Gender) %>%
  summarise(count = n())

print(Grouped)
```
```{r}
#Get the total education count for each level
Total_Education_Count <- employee %>%
  group_by(Education) %>%
  summarise(Total_Count = n())

print(Total_Education_Count)
```

```{r}
# Find the percentage of females in each level
percentage_of_female <- Grouped %>%
  filter(Gender == "Female") %>%
  left_join(Total_Education_Count, by = "Education") %>%
  mutate(percentage = (count / Total_Count) * 100) %>%
  select(Education, percentage)

print(percentage_of_female)
```

#(b)
### On a single plot, make three line plots, one for each City, with the variable JoiningYear on the x-axis and cumulative number of hires up to that year on the y-axis (Note: Each row is considered a hire).

```{r}
#Group employee dataframe by city and joining year and find the cumulative sum of hires by year
cumulative_hires_by_year <- employee %>%
  group_by(City, JoiningYear) %>%
  summarise(Cumulative_Hires = n()) %>%
  group_by(City) %>%
  mutate(Cumulative_Hires = cumsum(Cumulative_Hires))

cumulative_hires_by_year
```

```{r}
# Create a line plot for each City
ggplot(cumulative_hires_by_year, aes(x = JoiningYear, y = Cumulative_Hires, color = City)) +
  geom_line() +
  labs(x = "Joining Year", y = "Cumulative Number of Hires") +
  ggtitle("Cumulative Number of Hires by Joining Year")
```

#(c)
### Create side-by-side boxplots of Age vs Gender faceted by City on the rows and Education on the columns.

```{r}
# Create a side-by-side boxplot
ggplot(employee, aes(x = Gender, y = Age)) +
  geom_boxplot() +
  facet_grid(City ~ Education) +
  labs(x = "Gender", y = "Age") +
  ggtitle("Age vs Gender Faceted by City and Education")
```

#(d)
### For each JoiningYear, compute the ratio of the number of Bachelor’s degrees hired to Ph.D’s hired. Which year has the highest ratio and what is that ratio? For each City, compute the ratio of the number of Bachelor’s degrees hired to Ph.D’s hired. Which city has the highest ratio and what is that ratio? You have now found a year and a city. How many master’s level people were hired in the year and city that you just found?

```{r}
# Calculate the ratio of Bachelor's to Ph.D's for each JoiningYear
ratio1 <- employee %>%
  filter(Education %in% c("Bachelors", "PHD")) %>%
  group_by(JoiningYear, Education) %>%
  summarise(Count = n()) %>%
  pivot_wider(names_from = Education, values_from = Count, values_fill = 0) %>%
  mutate(Ratio = Bachelors / PHD) %>%
  arrange(desc(Ratio))

ratio1
```
###Which year has the highest ratio and what is that ratio?
###From the output, the year with the highest ratio is 2012 with a ratio of 31.5


```{r}
# Calculate the ratio of Bachelors to PHD for each City
ratio2 <- employee %>%
  filter(Education %in% c("Bachelors", "PHD")) %>%
  group_by(City, Education) %>%
  summarise(Count = n()) %>%
  pivot_wider(names_from = Education, values_from = Count, values_fill = 0) %>%
  mutate(Ratio = Bachelors / PHD) %>%
  arrange(desc(Ratio)) 
  
ratio2
```

### Which city has the highest ratio and what is that ratio?
###From the output, the city with the highest ratio is Pune, with a ratio of 42.17


```{r}
# How many master's level people were hired in the year and city with the highest ratio
masters_hired <- sum(employee$Education == "Masters" & employee$JoiningYear == 2012 & employee$City == "Pune")

cat("The number of masters level people from Pune and hired in 2012 is :", masters_hired)
```

#Question 2


```{r}
#Loading the dataset
stops <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI401/main/data/CTA_List_of_L_Stops.csv")


entries <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI401/main/data/CTA_entries.csv")
```

```{r}
#Reading the first few rows of the dataset
head(stops)
```
```{r}
#Getting the shape of stops
dim(stops)
```
```{r}
#Getting the columns
names(stops)
```
```{r}
#Reading the first few rows of the dataset
head(entries)
```
```{r}
#Getting the shape of entries
dim(entries)
```
```{r}
#Getting the columns
names(entries)
```
#(a)
### Using the stops data set, create a data set that only includes stops on the red line and removes duplicated MAP IDs. Merge the stops onto the entries data set. (Note: station id in entries links to MAP ID in stops!)

```{r}
#Filter stops for the red line and remove duplicated MAP IDs
red_line_stops <- stops %>%
  filter(RED == "true") %>%
  distinct(MAP_ID)
```

```{r}
#Merge stops onto entries using the station_id (MAP_ID) as the key
merged_data <- entries %>%
  inner_join(red_line_stops, by = c("station_id" = "MAP_ID"))

#Print the merged dataset
head(merged_data)
```

#(b)
###Using this merged data set, compute the average number of rides per day for each station on the red line. Next compute the average number of rides per day for each station for each month (i.e.average rides per day at Loyola at January, average rides per day at Loyola in February, etc.).

```{r}
#Calculate the average rides per day for each station on the red line
average_rides_per_station_day <- merged_data %>%
  group_by(stationname) %>%
  summarise(average_rides_per_day = mean(rides, na.rm = TRUE))

head(average_rides_per_station_day)
```

```{r}
# Convert the 'date' column to DateTime
merged_data$date <- as.Date(merged_data$date, format = "%m/%d/%Y")

# Extract the month from the 'date' column and create a new 'MONTH' column
merged_data$MONTH <- format(merged_data$date, "%m")

# Group by 'stationname' and 'MONTH' and calculate the average rides per day for each month
average_rides_per_station_month <- merged_data %>%
  group_by(stationname, MONTH) %>%
  summarise(average_rides_per_day_month = mean(rides, na.rm = TRUE))

head(average_rides_per_station_month)
```

#(c)
###Using this two variables found in the previous part, for each station create a variable called ratio which is the ratio of rides per day per month divided by rides per day. (This is essentially a measure of how far away ridership in a given month gets away from the overall average at a stop with values larger than 1 indicating a sharp increase and values less than 1 a sharp decrease). Display the top ten largest ratios and comment on and reason you believe these particular stops and months are the top of the list.



```{r}
#Calculate the ratio of rides per day per month
ratio_data <- average_rides_per_station_month %>%
  left_join(average_rides_per_station_day, by = "stationname") %>%
  mutate(ratio = average_rides_per_day_month/ average_rides_per_day)

head(ratio_data)
```

```{r}
#Get the top ten largest ratios
top_10_ratios <- ratio_data %>%
  arrange(desc(ratio)) %>%
  head(10)

top_10_ratios
```
###Addison-North Main and Fullerton stops are close to Wrigley Field, home of the Chicago Cubs, which attracts large crowds during baseball season. 


#(d)
### Compute the standard deviation of this ratio for each station and find the stations with the 5 largest standard deviations. Using these five stations, a line plot (one line for each station) with month on the x-axis and ratio on the y-axis.

```{r}
#Calculate the standard deviation of the ratio for each station
standard_deviation_data <- ratio_data %>%
  group_by(stationname) %>%
  summarise(sd_ratio = sd(ratio, na.rm = TRUE))
```

```{r}
#Find the stations with the 5 largest standard deviations
top_5_stations <- standard_deviation_data %>%
  arrange(desc(sd_ratio)) %>%
  head(5)

#Filter the ratio_data for the top 5 stations
top_5_data <- ratio_data %>%
  filter(stationname %in% top_5_stations$stationname)
```


```{r}
#Create a line plot for each station
library(ggplot2)

ggplot(top_5_data, aes(x = MONTH, y = ratio, group = stationname, color = stationname)) +
  geom_line() +
  labs(title = "Ratio of Rides per Day per Month For Top 5 Stations",
       x = "Month",
       y = "Ratio") +
  theme_minimal()
```


#Python Link: https://colab.research.google.com/drive/1TpBiPp0-tsJOCROuE0ErtoS-vNKMsn5f#scrollTo=eyMhriK76hyI