---
title: "DSCI401_HW2"
author: "Bernard Ofori Boateng"
date: "2023-09-20"
---

# Question 1
```{r}
# loading the teams dataset
teams <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI401/main/data/Teams.csv")
```

```{r}
# loading the first five rows
head(teams)
```

```{r}
# getting the shape of the dataframe
dim(teams)
```

```{r}
names(teams)
```

```{r}
# loading the tidyverse
suppressMessages(library(tidyverse))
```


## (a)
```{r}
teams_new <- teams %>% filter(yearID >= 2000 & yearID <= 2009) %>% select(yearID, W, L)
head(teams_new)
```

## (b)
```{r}
# filtering the teams data for HR>=200 for the Chicago cubs
cubs_HR <- teams %>% filter(teamID == "CHN" & HR >= 200) 

# finding the number of unique years the cubs had HR>=200
total_years <- length(unique(cubs_HR$yearID))
cat("Number of years the Chicago Cubs hit at least 200 HRs:" ,total_years)

# finding the median of the wins
median_wins <- median(cubs_HR$W)
cat("\nThe median number of wins in the seasons the Chicago Cubs hit at least 200 HRs:" ,median_wins)

```
## (c)
```{r}
# creating the factor election based on the presidential terms
teams <- teams %>% mutate(election = factor((yearID-1788) %/% 4 + 1))
head(teams)
```
```{r}
# finding the home runs for each 4 year term
HR_term <- teams %>% group_by(election) %>% summarise(HR_total = sum(HR))
head(HR_term)
```
```{r}
# finding the term with the most home runs
HR_term_max <- HR_term %>% filter(HR_total == max(HR_total))
HR_term_max
```
### The presidential term with the most home runs is the 58th preseidential term with a total of 24076 home runs

## (d)
```{r}
# loading the ggplot2 package
library(ggplot2)

# removing observations where league ID is missing
teams_lgID_complete <- teams %>% filter(!is.na(lgID))

# line plot of total home runs per season
ggplot(teams_lgID_complete, aes(x = yearID, y = HR, color = lgID)) + geom_line() + 
  labs(
        title = "Total Home Runs Per Season Stratified by League",
                                  x = "Year",
                                  y = "Total Home Runs",
                                  color = "League ID"
                                )
```

## (e)
```{r}
# creating the indicator variable called winning record 
teams <- teams %>% mutate(winning_record = W > L)
head(teams)
```
```{r}
# scatter plot of Runs (R) vs Runs against (RA)
ggplot(teams, aes(x = R, y = RA, color = winning_record)) + geom_point() + 
  labs(
        title = "Scatter Plot of Runs (R) vs Runs against (RA)",
                                  x = "Runs (R)",
                                  y = "Runs Against (RA)",
                                  color = "Winning Record"
                                )
```


# Question 2

```{r}
library(mdsr)
# getting a glimpse of the dataset
glimpse(Violations)

```
```{r}
# loading the first 5 rows of the violations dataset
head(Violations)
```

## (a)
```{r}
# proportion of inspections graded into A and not A for each boro 

boro_proportion <- Violations %>%
  mutate(grade_A_or_NotA = ifelse(is.na(grade) | grade != "A", "Not_A", "A")) %>%
  group_by(boro, grade_A_or_NotA) %>%
  summarise(observation_count = n()) %>%
  mutate(proportion_A = sum(observation_count[grade_A_or_NotA == "A"] / sum(observation_count))) %>%
  filter(grade_A_or_NotA == "A") 

boro_proportion <- boro_proportion %>%
  select(boro, proportion_A)

boro_proportion
```

## (b)

```{r}
# finding the top ten dba's with the most no. of inspections
top_dba <- Violations %>%
  group_by(dba) %>%
  summarise(no.of_inspections = n(),mean_score = mean(score,na.rm=TRUE)) %>%
  arrange(desc(no.of_inspections))
top10_dba <- top_dba[1:10,] %>%
  arrange(mean_score) %>%
  select(dba, mean_score)
top10_dba
```
### STARBUCKS had the lowest average inspection score



## (c)
```{r}
# getting data from Manhattan zip codes with 50 or more inspections
median_violation_score <- Violations %>% 
  filter(boro == "MANHATTAN") %>%
  group_by(zipcode) %>%
  filter(n() >= 50) %>%
  summarise(median_score = median(score, na.rm = TRUE),no_of_inspections = n()) %>% 
  arrange(zipcode)
median_violation_score
```
### From observations, there appears to be an inverse relationship between the median score and number of inspections. The zip codes that had more inspections had lower median score. This could be due that the zip codes with more inspections complied with regulations. 



# Python Link: https://colab.research.google.com/drive/1qV6zAsS0Mw95pVad3ZnYWL1S32FgTbhw#scrollTo=6eOzEthQgg7l&uniqifier=1

