---
title: "DSCI401_HW3"
author: "Bernard Ofori Boateng"
date: "2023-10-02"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Question 1
```{r}
# loading the package and tables
library(tidyverse)
library(Lahman)
library(mdsr)
head(People)
head(Batting)
head(Pitching)
```

# (a)

### (10 points) Name every player in baseball history who has accumulated at least 300 home runs (HR) AND at least 300 stolen bases (SB). You can find the first and last name of the player in the People data frame. Join this to your result along with the total home runs and total bases stolen for each of these elite players.

```{r}
# lets get the total HR and SB for each player and filter for >= 300
elite_totals <- Batting %>% 
  select(playerID, HR, SB) %>%
  group_by(playerID) %>%
  summarise(total_HR = sum(HR), total_SB = sum(SB)) %>%
  filter(total_HR >= 300, total_SB >= 300)
  
# join totals with People table to get player names
elite_totals <- elite_totals %>% left_join(People, by = c("playerID" = "playerID")) %>%
  select(playerID, nameFirst, nameLast, total_HR, total_SB) 

elite_totals
```

# (b)

### (10 points) Similarly, name every pitcher in baseball history who has accumulated at least 300 wins (W) and at least 3,000 strikeouts (SO).

```{r}
elite_pitchers <- Pitching %>% 
  select(playerID, W, SO) %>%
  group_by(playerID) %>%
  summarise(total_W = sum(W), total_SO = sum(SO)) %>%
  filter(total_W >= 300, total_SO >= 3000) %>%
  left_join(People, by = c("playerID" = "playerID")) %>%
  select(playerID, nameFirst, nameLast, total_W, total_SO)

elite_pitchers
```

# (c)

### (10 points) Identify the name and year of every player who has hit at least 50 home runs in a single season. Which player had the lowest batting average in that season?

```{r}
# Identify the name and year of every player who has hit at least 50 home runs in a single season.
part_a <- Batting %>%
  select(playerID, yearID, HR) %>%
  group_by(playerID, yearID) %>%
  summarise(HR_total = sum(HR)) %>%
  filter(HR_total >= 50) %>%
  left_join(People, by = c("playerID" = "playerID")) %>%
  select(playerID, yearID, nameFirst, nameLast, HR_total) %>%
  arrange(desc(HR_total))
  
part_a  
```

```{r}
# join players to Batting dataset
part_b <- part_a %>%
  inner_join(Batting, by = c("playerID" = "playerID" , "yearID" = "yearID"))
part_b
```

```{r}
# find the lowest batting average
part_c <- part_b %>% 
  mutate(batting_avg = H/AB) %>%
  select(playerID, yearID, nameFirst, nameLast, batting_avg) %>%
  arrange(batting_avg)

part_c
```
## Mark McGwire had the lowest batting average in that season with a batting average of 0.252874


# Question 2

```{r}
library(nycflights13)
head(flights)
```
```{r}
head(planes)
```

# (a)

### (10 points) What is the oldest plane (specified by the tailnum variable) that flew from New York City airports in 2013?
```{r}
oldest_plane <- flights %>%
  select(tailnum, dep_time) %>%
  drop_na(dep_time)

oldest_plane %>%
  inner_join(planes, by = c("tailnum" = "tailnum")) %>%
  drop_na(year) %>%
  arrange(year) %>%
  head(5)
```
## The oldest plane by tailnum is N381AA


# (b)

### (10 points) How many airplanes that flew from New York City are included in the planes table?

```{r}
# find the number of unique tailnum in the flights table
flights_1 <- oldest_plane %>%
  distinct(tailnum) 

flights_2 <- planes %>%
  inner_join(flights_1, by = c("tailnum" = "tailnum")) %>%
  nrow()
flights_2
```

```{r}
num_planes <- planes %>%
  distinct(tailnum) %>%
  nrow()
num_planes
```
### 3316 airplanes that flew from NYC are included in planes (3322) table.


# Question 3

```{r}
dat <- data.frame(grp = c("A","A","B","B"),
sex = c("F","M","F","M"),
meanL = c(0.225,0.47,0.325,0.547),
sdL = c(0.106,.325,.106,.308),
meanR = c(.34,.57,.4,.647),
sdR = c(0.0849, 0.325, 0.0707, 0.274)
)

dat
```
# (a)

### Generate the code to convert the following data frame to wide format.
```{r}
# use pivot_wider to convert the data frame to wide format
dat_wide <- pivot_wider(
  data = dat,
  names_from = sex,
  values_from = c(meanL, sdL, meanR, sdR)
)

dat_wide
```


# Question 4
```{r}
library(pccc)
head(pccc_icd10_dataset)

```

# (a)

### Remove all the columns labeled with ”g” and a number.

```{r}
# Remove columns labeled with "g" followed by a number
pccc_icd10_dataset_new <- pccc_icd10_dataset %>%
  select(-contains("g"))

names(pccc_icd10_dataset_new)
```

# (b)

### Convert this to a long data set with three columns: id, type (pc or dx), and code.

```{r}
# Create a long data set
pccc_icd10_dataset_long <- pccc_icd10_dataset_new %>%
  pivot_longer(
    cols = -id, 
    names_to = "type", 
    values_to = "code")

# Print the data frame
head(pccc_icd10_dataset_long)


```


## Python Link: https://colab.research.google.com/drive/1A756QiI9M48xz3J2JGW9GMw7GTelum5g#scrollTo=y5nZkrVD6Fpl&uniqifier=3